Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-04-14 paper/main.tex
hamming v0
Update template, fix affiliation errors
fit in 8 pages + ref

\author{Ekhiñe Irurozki}
\affiliation{
  \institution{Basque Center for Applied Mathematics}
}
  \email{eirurozki@bcamath.org}

\author{Manuel López-Ibáñez}
\orcid{0000-0001-9974-1295}
\affiliation{
  \institution{University of Málaga}
  \city{Málaga}
  \state{Spain}
}
\email{manuel.lopez-ibanez@uma.es}

@@

This paper is structured as follows. Section~\ref{sec:backgroud} summarizes fundamental concepts about permutation spaces, uBorda, black-box combinatorial optimization under a limited budget and CEGO, a Bayesian optimizer for such problems. Section~\ref{sec:umm} presents the UMM algorithm, which is the main contribution of this paper. UMM is analyzed using the experimental setup presented in Section~\ref{sec:setup} and results are presented in Section~\ref{sec:analysis}. We end the paper with overall conclusions and a call for further research on this type of optimization scenario (Section~\ref{sec:conclusions}).

@@

Permutations are defined as bijections of the set $[n]$ of integers onto itself. The set of all permutations of $n$ values is denoted as $S_n$ and has cardinality $n!$. We will use the one-line notation
% \MANUEL{what is one-line?}
and denote permutations with Greek letters, $\sigma  = (\sigma(1), \ldots , \sigma(n))$. We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi = (\sigma(\pi(1)), \ldots, \sigma(\pi(n)))$
% \MANUEL{Is this correct?}
and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma^{-1}=(1, 2, 3, \dotsc,n)$ always holds.

@@

The learning rate is
controlled by parameter $0 < \rho < 1$. 
% \MANUEL{Don't we have to say something about $f(\sigma) > 1$?}
% \EKHINE{es decreciente igual}
In UMM, the value of $\rho$ depends on the set of true fitness evaluations saved in $S$
and it is set at each iteration such that the largest $100r_1$\% of the mass of
the weights is concentrated in the best $100r_2$\% of the solutions in $S$:
%
% 
\begin{equation}
%  \begin{split}
    r_1 \cdot \sum_{\tuple{\sigma,f(\sigma)}\in S}\rho^{\text{scale}(f(\sigma))} =  \sum_{\tuple{\sigma', f(\sigma')}\in S'}\rho^{\text{scale}(f(\sigma'))}  %\\
  %    & \text{where}\;|S'| = r_2|S| \land \forall \sigma'\in S', \sigma\in S\setminus S' : f(\sigma')\leq f(\sigma).
%  \end{split}
\end{equation}
%
where $|S'| = r_2|S| \land \forall \tuple{\sigma',f(\sigma')}\in S', \tuple{\sigma,f(\sigma)}\in S\setminus S' : f(\sigma')\leq f(\sigma)$, that is, $S'$ contains the $r_2|S|$ elements of $S$ with the best fitness values. 
%The values of $r_1, r_2 \in (0, 1)$ are user-defined and control the balance between exploration and intensification. %: extreme different values bias the sampling towards the best solutions found so far, which may lead to quick stagnation, whereas values (0.5,0.5) make the sampling more uniform, thus slowing down the learning.

@@

\begin{figure}[tb]
  \centering%
  \smaller
\begin{minipage}{0.49\linewidth}
  \centering
    \texttt{N-t59d11xx} (LOP)\\[-0.2ex]
    \includegraphics[width=\textwidth]{../img/heatmap_lop_IO_N-t59d11xx}
  \end{minipage}
  \begin{minipage}{0.49\linewidth}
    \centering
    \texttt{N-sgb75.02} (LOP)\\[-0.2ex]
 \includegraphics[width=\textwidth]{../img/heatmap_lop_SGB_N-sgb75_02}
\end{minipage}\\\vskip 1em
\begin{minipage}{0.49\linewidth}
  \centering
  \texttt{reC13} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/heatmap_pfsp_rec13_txt}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \centering
  \texttt{reC19} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/heatmap_pfsp_rec19_txt}
\end{minipage}
% \begin{minipage}{0.49\linewidth}
%   \centering
%   \texttt{kra32} (QAP)\\[-0.5ex]
% \includegraphics[width=\textwidth]{../img/heatmap_kra32_dat}
% \end{minipage}
\caption{Mean fitness (over 10 runs) of different configurations of $r_1$ and $r_2$ (darker is better).\label{fig:heatmaps}}
\end{figure}

>>>>

\author{Ekhine Irurozki}
\orcid{0000-0003-3218-5735}
\affiliation{
  \institution{Telecom Paris}
  \city{Paris}
  \country{France}
}
\additionalaffiliation{Basque Center for Applied Mathematics}

\email{irurozki@telecom-paris.fr}

\author{Manuel López-Ibáñez}
\orcid{0000-0001-9974-1295}
\affiliation{
  \institution{University of Málaga}
  \city{Málaga}
  \country{Spain}
}
\additionalaffiliation{University of Manchester, UK}
\email{manuel.lopez-ibanez@uma.es}

@@

This paper is structured as follows. Section~\ref{sec:backgroud} summarizes fundamental concepts.
% about permutation spaces, uBorda, black-box combinatorial optimization under a limited budget and CEGO, a Bayesian optimizer for such problems.
Section~\ref{sec:umm} presents the UMM algorithm, which is the main contribution of this paper. UMM is analyzed using the experimental setup presented in Section~\ref{sec:setup} and results are presented in Section~\ref{sec:analysis}. We end the paper with overall conclusions and a call for further research on this type of optimization scenario (Section~\ref{sec:conclusions}).

@@

Permutations are %defined as
bijections of the set $[n]$ of integers onto itself. The set of all permutations of $n$ values is denoted by $S_n$ and has cardinality $n!$. We will use the one-line notation
% \MANUEL{what is one-line?}
and denote permutations with Greek letters, $\sigma  = (\sigma(1), \ldots , \sigma(n))$. The composition of $\sigma$ and $\pi$ is $\sigma\pi = (\sigma(\pi(1)), \ldots, \sigma(\pi(n)))$
% \MANUEL{Is this correct?}
and the inverse of $\sigma$ is $\sigma^{-1}$, for which the relation $\sigma\sigma^{-1}=(1, 2, 3, \dotsc,n)$ always holds.

@@

The learning rate is
controlled by parameter $0 < \rho < 1$. 
% \MANUEL{Don't we have to say something about $f(\sigma) > 1$?}
% \EKHINE{es decreciente igual}
In UMM, the value of $\rho$ depends on the set of true fitness evaluations saved in $S$
and it is set at each iteration such that the largest $100r_1$\% of the mass of
the weights is concentrated in the best $100r_2$\% of the solutions in $S$:
%
% 
\begin{equation}
%  \begin{split}
    r_1 \cdot \sum_{\tuple{\sigma,f(\sigma)}\in S}\rho^{\text{scale}(f(\sigma))} =  \sum_{\tuple{\sigma', f(\sigma')}\in S'}\rho^{\text{scale}(f(\sigma'))}  %\\
  %    & \text{where}\;|S'| = r_2|S| \land \forall \sigma'\in S', \sigma\in S\setminus S' : f(\sigma')\leq f(\sigma).
%  \end{split}
\end{equation}
%
where $|S'| = r_2|S| \land \forall \tuple{\sigma',f(\sigma')}\in S', \tuple{\sigma,f(\sigma)}\in S\setminus S' : f(\sigma')\leq f(\sigma)$, i.e., $S'$ contains the $r_2|S|$ elements of $S$ with the best fitness. 
%The values of $r_1, r_2 \in (0, 1)$ are user-defined and control the balance between exploration and intensification. %: extreme different values bias the sampling towards the best solutions found so far, which may lead to quick stagnation, whereas values (0.5,0.5) make the sampling more uniform, thus slowing down the learning.

@@

\begin{figure}[!tb]
  \centering%
  \smaller
\begin{minipage}{0.49\linewidth}
  \centering
    \texttt{N-t59d11xx} (LOP)\\[-0.2ex]
    \includegraphics[width=\textwidth]{../img/heatmap_lop_IO_N-t59d11xx}
  \end{minipage}
  \begin{minipage}{0.49\linewidth}
    \centering
    \texttt{N-sgb75.02} (LOP)\\[-0.2ex]
 \includegraphics[width=\textwidth]{../img/heatmap_lop_SGB_N-sgb75_02}
\end{minipage}\\\vskip 1em
\begin{minipage}{0.49\linewidth}
  \centering
  \texttt{reC13} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/heatmap_pfsp_rec13_txt}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \centering
  \texttt{reC19} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/heatmap_pfsp_rec19_txt}
\end{minipage}
% \begin{minipage}{0.49\linewidth}
%   \centering
%   \texttt{kra32} (QAP)\\[-0.5ex]
% \includegraphics[width=\textwidth]{../img/heatmap_kra32_dat}
% \end{minipage}
\caption{Mean fitness (over 10 runs) of different configurations of $r_1$ and $r_2$ (darker is better).\label{fig:heatmaps}}
\end{figure}
